Cover letter for the original IJRR submission

### 2017-10-03 reviews_tro/reviewer_response.tex

\quot{
Also, where you describe "sensitivity" vs "specificity", it would be
useful to give a more intuitive explanation, example [\dots]}

\as{We updated Section VIII.B.1) to illustrate the meaning of those terms.}
\qt{We then computed the sensitivity and specificity of the reachability condition.  In this context, the sensitivity refers to the percentage of configurations in $C_{Reach}^0$, effectively belonging to \gls{$C_{Contact}^0$}. If a sampled configuration is in $C_{Reach}^0$, but our method is unable to generate a contact configuration from it, as a result the sensitivity
decreases. The sensitivity thus illustrates the confidence we have that any configuration in $C_{Reach}^0$ will effectively lead to a contact configuration.
Conversely, the specificity refers to the percentage of configurations \textbf{not} in $C_{Reach}^0$, effectively \textbf{not} belonging to \gls{$C_{Contact}^0$}.
If a sampled configuration is \textbf{not} in $C_{Reach}^0$, but our method is able to generate a contact configuration from it, as a result the specificity
decreases. The specificity thus illustrates the confidence we have that all configurations that allow contact creation belong to $C_{Reach}^0$ (or informally, the confidence that we are not missing valid solutions).
We thus look for a compromise between sensitivity and specificity.} \done

\quot{
However, I suggest the paper will be improved if
the presentation focuses more on highlighting (a) what is novel, and
(b)
what take-away messages you have, about why the whole framework is
designed the way it is [within the main text; not hidden in
appendices].}

\as{Thanks for the comments. We believe our new version of the papers answers to a). In particular, as mentionned above the novel content have been moved from the Appendix
to the core of the paper. Regarding b), the take-away message is that to address the hard problem of multi contact planning, it is necessary to use 
abstraction and simplification, through the decomposition of the problem into three sub-problems, and through dimensionality reduction.
We believe that the paper develops these aspects to explain the design of this planner. In the introduction, it is done
with an extensive state of the art that explains how the community came up with the P1, P2, P3 decomposition to tackle a problem that is too hard. In the conclusion, it is explained 
with an analysis of the results we obtain, that justify our pragmatic choice of going further in the abstraction to obtain interactive planning performances. To follow the reviewer's advice, we updated the conclusion
to include a synthesized give away message.}
\qt{In this paper we consider the multi-contact planning problem, formulated as three sub-problems  $\mathcal{P}_1$,  $\mathcal{P}_2$,  $\mathcal{P}_3$, addressed sequentially. While we propose a global framework that handles all these problems, our contribution focuses on  $\mathcal{P}_1$ and  $\mathcal{P}_2$.
The first problem $\mathcal{P}_1$ consists in computing an \gls{equilibrium feasible} guide path for the root of the robot;
the second problem $\mathcal{P}_2$ is the computation of a discrete sequence of whole-body configurations along the root path.
We believe that this decomposition is currently the most promising approach towards
a global resolution of the problem. We also claim to have achieved a significant step towards this objective thanks to the dimensionality reduction provided by
the reachability condition. With our results and the release of our source code, we hope to inspire further research in this direction.}\done

@@

\quot{
p.6 – this would be a great place to give more ``intuition'' for the heuristics. Hw seems to be used in essentially all foot contact, with $h_{efort}$ for hand/arm contacts, and ALSO $h_{vel}$ not seeming to be used at all.}
\as{Thanks for the suggestion. The text was updated accordingly, and the heuristic $h_{vel}$ was removed since indeed it is not used.} 
\qt{In all our experiments, the heuristic $h$ is implemented as a variation of a manipulability-based heuristic [27]. The manipulability is a real number that quantifies how 
``good'' a configuration is to perform a given task, based on the analysis of the Jacobian matrix. With such heuristics, a configuration can be chosen because it is far from singularities, and thus allows mobility in all directions. On the contrary, it can be chosen because it is particularly efficient to exert a force in a desired direction. In our experiments, the former solution is usually chosen for computing leg contacts, while the latter is used for computing hand contacts. We recall the manipulability measure and its derivatives in Appendix B.}\done

\quot{
p.6 six fingered hand is slightly misleading, since no rolling (as fully admitted in the text). I'd just cut this? \\
p.9 Not sure if Fig. 14 is worth including. It looks like its just Fig. 9 from the ISRR paper (?) [20], and there is still no ``full solution'' here -
 just an illustration of potential future work. 
 (At a minimum, the figure/solution should probably be referenced as having already appeared in the previous ISSR paper?)
}

\as{As also suggested by reviewer 4, we removed the references to the results of the previous work [20], including those.}\done

@@

\quot{github link seems to be dead? (The link on p. 15 is OK\dots should they be the same?)}

\as{We are suprised by this comment because after testing it, the link seems to be valid. In any case, in accordance with reviewer 6, we changed all the links source codes to provide a SHA link that will bring the user to the accurate revision of the projects considered.}
\qt{(Section VII) Both HPP and our planner can be simply downloaded and compiled by following the instructions on
\url{https://humanoid-path-planner.github.io/hpp-doc/download.html?branch=rbprm}. \\
(Appendix C) To address the other scenarios, we propose a new implementation, entirely open source (\url{https://github.com/stonneau/python_sandbox/releases/tag/tro_paper} ), which can be integrated directly in our motion planner software. }\done

\quot{p.5, on right, ``less than epsilon'' between end effector and world - but doesn’t the orientation of the end effector matter tremendously here, too?}
\as{
The orientation of the end effector matters, but is not problematic regarding the way we generate contacts.
Using an epsilon is an effective way to remove many unrelevant configurations, and
we believe that more complex solutions aiming at taking into account the orientation would end up being counter productive.
 Indeed, using only 3D distance, we will end up with a few configurations
that can't be projected correctly, and will end up being discarded. But the majority of irrelevant configurations would already have been rejected with
the efficient 3D intersection query. Conversely, introducing a 6D octree rather than the current 3D one would make the problem much more complex: instead
of using well known 3D operations to retrieve our candidate configurations, we would have to implement our own, new solutions. The added cost of doing this is probably 
much higher than the computational cost of discarding a few configurations that could not be projected.  } \done

\quot{remaining remarks by reviewer 1}
\as{The remaining comments from reviewer 1 were all addressed. Because they did not call for a discussion, as they are essentially related to style or typos, we do not list them here.}\done

@@

\quot{* I am not sure why the $C_{reach}$ hypothesis is described as the ``strong''
hypothesis in, e.g., Sect. IV. Is there a corresponding weak
hypothesis?}
\as{We apologize for the confusion, that comes from an inaccurate use of English. What we meant by ``strong'' is that the hypothesis is constraining, because
it defines a solution space different than the real one, and that we need to demonstrate that it makes sense.  We removed ``strong'' and replaced it by ``main'' to clarify. }\done

\quot{* In Sect. IV.B, the authors state that the ideal shape $B^*$ ``has no
explicit definition''. For a particular configuration  $q_0$, $B^*$ is simply
the surface of the robot, right? So I assume that this claim refers to
there being no explicit definition of B* which is the volume around
$B^*(q_0)$ for every $q_0$? I think I agree with the authors, but I found
this claim hard to follow and would appreciate some clarification}
\as{We updated the Section IV B. to better explain our claim. Indeed, as the reviewer states $B^*$  is the volume for every $q_0$. If such a volume existed, we would know exactly whether or not a given root configuration can result in a valid contact configuration.
We of course doubt that $B^*$ actually exists. }
\qt{The ideal shape $B^*, W^0 \subset B^* \subset B^{\textrm{\it Suf}}$ would define a necessary \textbf{and} sufficient condition for contact creation. 
It would guarantee that any root configuration $\mathbf{q}^{0} \in B^*$ would result in a contact configuration, while any $\mathbf{q}^{0} \notin B^*$ could not.
To our knowledge  $B^*$ has no explicit definition.
Therefore, we approximate $B^*$ to define the contact reachable space $C_{Reach}^0$.}\done

\quot{* In Sect. V.A., the authors mention that the ability to plan a contact
break and creation with the same limb allows the creation of dynamic
motions. This is quite an important ability for an otherwise
quasi-static planner, but this ability isn't described again until the
appendix. It would be worth explaining why this is true in the paper
itself. }
\as{Thanks for the suggestion. We added a paragraph to better explain this in Section V.A.}
\qt{This representation is sufficient to describe all the contact phases, because the single support phase is implicitely described. Furthermore it removes the need to compute a single-support quasi-static configuration as in the example. Indeed, there might be a case where no quasi-static solution exists for the single support phase (because of the environment), but
there exists a dynamic motion connecting the two double support states. Such motion will be computed by our framework, because the quasi-static constraint is only required at the contact planning phase; as shown in the companion video, and explained in Appendix C, our framework is able to produce dynamic motions.
}
\done

### 2017-10-03 body/conclusion.tex
most modifs included, missing pdf of reviewer 1 minor comments, sota, and some others
update on contact break / creation

Another source of computational cost identified in previous works is the verification of \equilibriumfeasibility. 
The strong assumption of our work is that for the class of problems we consider \contactreachability\ implies \equilibriumfeasibility.
Our scenarios show that the  assumption is verified in the majority of cases when at least one contact surface is \textit{quasi flat}~\cite{Prete2016}, that is when
the friction cone of the contact surface contains the direction opposite to the gravity.
Figure~\ref{fig:dedefeas} illustrates this observation, demonstrated empirically by the third column of Table~\ref{tab:requestpercent}. In the worst case, in our experiment
the assumption was verified for 82\% of the total amount of trials that verified \textit{contact reachability}.
In the example of \cite{Bouyarmane2009}, the verification of \equilibriumfeasibility\ implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$, requiring
several minutes of planning. Our method, in comparison, takes from a few milliseconds to several seconds.

@@

In this paper we consider the multi-contact planning problem, formulated as two sub-problems that we address sequentially.
The first problem $\mathcal{P}_1$ consists in computing an \gls{equilibrium feasible} guide path for the root of the robot;
 the second problem $\mathcal{P}_2$ is the computation of a discrete sequence of whole-body configurations along the root path.

@@

Our results demonstrate that our method allows a pragmatic compromise between three 
criteria that are hard to reconcile: generality, performance, and quality of the solution, making it the first acyclic contact
planner compatible with \gls{interactive} applications.
%
\textbf{Regarding generality}, the \textit{reachability condition}, coupled with an approach based on limb decomposition, 
allows the method to address automatically arbitrary legged robots.
%
\textbf{Regarding performance}, our framework is efficient in addressing both \Pa and $\mathcal{P}_2$. This results in \gls{interactive} computation times.
%
\textbf{Regarding the quality of the paths}, we are able to compute
\gls{equilibrium feasible} paths in all the presented scenarios, with high success rates.
As for \cite{Bouyarmane2009}, failures can still occur, due to the approximate condition used to compute the guide path.
The low computational burden of our framework however allows for fast re-planning in case of failure.
Furthermore, because of this approximation, the guide search is not complete. The choice is deliberate, because we believe
that it is necessary to trade completeness for efficiency at all stages of the planner.
However, one direction for future work is to focus on a more accurate formulation of $C_{reach}^0$ to improve
the approximation.

>>>>

Another source of computational cost identified in previous works is the verification of \equilibriumfeasibility. 
The main assumption of our work is that for the class of problems we consider \contactreachability\ implies \equilibriumfeasibility.
Our scenarios show that the  assumption is verified in the majority of cases when at least one contact surface is \textit{quasi flat}~\cite{Prete2016}, that is when
the friction cone of the contact surface contains the direction opposite to the gravity.
Figure~\ref{fig:dedefeas} illustrates this observation, demonstrated empirically by the third column of Table~\ref{tab:requestpercent}. In the worst case, in our experiment
the assumption was verified for 82\% of the total amount of trials that verified \textit{contact reachability}.
In the example of \cite{Bouyarmane2009}, the verification of \equilibriumfeasibility\ implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$, requiring
several minutes of planning. Our method, in comparison, takes from a few milliseconds to several seconds.

@@

In this paper we consider the multi-contact planning problem, formulated as three sub-problems  $\mathcal{P}_1$,  $\mathcal{P}_2$,  $\mathcal{P}_3$, addressed sequentially. While we propose a global framework that handles all these problems, our contribution focuses on  $\mathcal{P}_1$ and  $\mathcal{P}_2$.
The first problem $\mathcal{P}_1$ consists in computing an \gls{equilibrium feasible} guide path for the root of the robot;
the second problem $\mathcal{P}_2$ is the computation of a discrete sequence of whole-body configurations along the root path.
We believe that this decomposition is currently the most promising approach towards
a global resolution of the problem. We also claim to have achieved a significant step towards this objective thanks to the dimensionality reduction provided by
the reachability condition. With our results and the release of our source code, we hope to inspire further research in this direction.

@@

Our results demonstrate that our method allows a pragmatic compromise between three 
criteria that are hard to reconcile: generality, performance, and quality of the solution, making it the first acyclic contact
planner compatible with \gls{interactive} applications.

%
\textbf{Regarding generality}, the \textit{reachability condition}, coupled with an approach based on limb decomposition, 
allows the method to address automatically arbitrary legged robots.
%
\textbf{Regarding performance}, our framework is efficient in addressing both \Pa and $\mathcal{P}_2$. This results in \gls{interactive} computation times.
%
\textbf{Regarding the quality of the paths}, we are able to compute
\gls{equilibrium feasible} paths in all the presented scenarios, with high success rates.
As for \cite{Bouyarmane2009}, failures can still occur, due to the approximate condition used to compute the guide path.
The low computational burden of our framework however allows for fast re-planning in case of failure.
Furthermore, because of this approximation, the guide search is not complete. The choice is deliberate, because we believe
that it is necessary to trade completeness for efficiency at all stages of the planner.
However, one direction for future work is to focus on a more accurate formulation of $C_{reach}^0$ to improve
the approximation.

### 2017-10-03 body/contact.tex
update on contact break / creation

\subsection{Definition of a contact sequence}
In previous contributions~\citep{DBLP:conf/iser/EscandeKMG08}, a contact plan is defined as a sequence of quasi-static equilibrium configurations
for each contact phase. For instance, a walk cycle would be described by three key configurations: a double-support configuration, a single-support configuration (a contact is broken), and another double-support configuration (a contact is created). 
Our definition of contact plan differs: between two consecutive configurations we allow both a contact break and a contact creation---if they are on the same effector. 
In the previous example, our contact plan would simply consist of the two double-support configurations. 
%However, if the contact broken and the contact created sequentially are not on the same effector, our planner outputs two distinct states.
This representation is sufficient to describe all the contact phases. Furthermore it removes the need to have single-support quasi-static configurations as in the example. 
As shown in the companion video, this allows our framework to produce dynamic motions.

>>>>

\subsection{Definition of a contact sequence}
In previous contributions~\citep{DBLP:conf/iser/EscandeKMG08}, a contact plan is defined as a sequence of quasi-static equilibrium configurations
for each contact phase. For instance, a walk cycle would be described by three key configurations: a double-support configuration, a single-support configuration (a contact is broken), and another double-support configuration (a contact is created). 
Our definition of contact plan differs: between two consecutive configurations we allow both a contact break and a contact creation---if they are on the same effector. 
In the previous example, our contact plan would simply consist of the two double-support configurations. 
%However, if the contact broken and the contact created sequentially are not on the same effector, our planner outputs two distinct states.
This representation is sufficient to describe all the contact phases, because the single support phase is implicitely described. Furthermore it removes the need to compute a single-support quasi-static configuration as in the example. Indeed, there might be a case where no quasi-static solution exists for the single support phase (because of the environment), but
there exists a dynamic motion connecting the two double support states. Such motion will be computed by our framework, because the quasi-static constraint is only required at the contact planning phase; as shown in the companion video, and explained in Appendix~\ref{app:optim}, our framework is able to produce dynamic motions.
